---
title: "ATP Predictor — Exploratory Data Analysis"
author: "Mona Ascha"
format:
  html:
    toc: true
    code-fold: true
    df-print: kable
execute:
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(ggplot2)
library(here)

# EDA only -- the pipeline itself lives in tested R/ functions. This
# document SOURCES them; it does not redefine any logic.
source(here::here("R", "impute_heights.R"))
source(here::here("R", "load_matches.R"))

# Raw cleaned matches (for surface / height / who-wins EDA) ...
matches <- clean_matches(load_atp_matches(2020:2026))

# ... and the leak-free modelling frame produced by
# analysis/build_training_data.R (run that first if these are stale).
train <- read.csv(here::here("data", "data_train.csv"), check.names = FALSE)
ids   <- read.csv(here::here("data", "data_ids.csv"),   check.names = FALSE)
```

## Dataset at a glance

```{r overview}
d <- as.Date(as.character(ids$tourney_date_P_1), format = "%Y%m%d")
tibble::tibble(
  rows          = nrow(train),
  features      = ncol(train),
  date_from     = format(min(d)),
  date_to       = format(max(d)),
  label_balance = round(mean(ids$Win_P_1), 3)
)
```

The seeded 50/50 player-slot assignment is why the label balance sits
at ~0.5 — a model that ignores the features can do no better than a
coin flip, by construction.

## Who wins the most? (sanity / domain check)

```{r top-players}
winner <- ifelse(ids$Win_P_1 == 1, ids$name_P_1, ids$name_P_2)
as.data.frame(sort(table(winner), decreasing = TRUE)[1:15])
```

The usual elite names dominate — a basic sanity check that the
slot/label plumbing is correct.

## Win rate by surface

```{r surface}
matches %>%
  count(surface) %>%
  mutate(share = round(n / sum(n), 3)) %>%
  arrange(desc(n))
```

Hard courts dominate the calendar, then clay, then grass — consistent
with the ATP tour structure.

## Player height

```{r height}
ht <- c(matches$winner_ht, matches$loser_ht)
ht <- ht[!is.na(ht)]
ggplot(data.frame(ht = ht), aes(ht)) +
  geom_histogram(binwidth = 2, fill = "#0066FF", colour = "white") +
  labs(x = "height (cm)", y = "player-matches",
       title = "Height distribution") +
  theme_minimal()
```

## Feature signal is weak individually — by design, not by accident

```{r feature-corr}
y   <- ids$Win_P_1
num <- train[sapply(train, is.numeric)]
num <- num[, sapply(num, function(c) all(is.finite(c)))]   # skip Inf bp_ratio
cors <- sort(sapply(num, function(c) cor(c, y, use = "complete.obs")),
             decreasing = TRUE)
data.frame(top = round(head(cors, 8), 3))
```

No single feature is strongly correlated with the outcome (|r| well
under 0.25). This is the *expected* profile after the step-4 leakage
fixes — the model's accuracy comes from XGBoost combining ~100 weak
honest signals, not from one outcome-encoding feature.

## Elo is leak-free (the headline correction)

```{r elo-asof}
# If as-of Elo were still the old end-of-history value, each player's
# Elo would be a single constant -> within-player sd == 0.
chk <- data.frame(name = ids$name_P_1, elo = train$latest_elo_av_P_1) %>%
  group_by(name) %>% filter(n() >= 15) %>%
  summarise(sd_elo = sd(elo, na.rm = TRUE), .groups = "drop")
tibble::tibble(
  players          = nrow(chk),
  median_within_sd = round(median(chk$sd_elo, na.rm = TRUE), 2)
)
```

Median within-player Elo sd is well above zero: each player's rating
moves over time, confirming the as-of-match (leak-free) Elo. Before
the fix this value was exactly **0**.

## Why pooled accuracy is high but honest

```{r competitiveness}
elo_gap <- abs(train$latest_elo_av_P_1 - train$latest_elo_av_P_2)
data.frame(
  bucket = c("lopsided (|Elo gap| >= median)", "competitive (< median)"),
  n      = c(sum(elo_gap >= median(elo_gap)), sum(elo_gap < median(elo_gap)))
)
```

Many ATP matches are lopsided (elite vs. qualifier) and trivially
predictable; the honest difficulty signal is accuracy on the
*competitive* subset (~0.76), not the pooled figure (~0.86). See
`README.md` for the full model performance table.
